---
title: "Lab 2: Introduction to data"
author: "Sanielle Worrell"
output: html_document
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(openintro)
library(tinytex)
```

Load flight data

```{r load-data}
data(nycflights)
```

## Analysis

### Departure delays

Let's start by examining the distribution of departure delays of all flights with a histogram.

```{r hist-dep-delay}
ggplot(data = nycflights, aes(x = dep_delay)) +
  geom_histogram()
```

```{r hist-dep-delay-bins}
ggplot(data = nycflights, aes(x = dep_delay)) +
  geom_histogram(binwidth = 15)
ggplot(data = nycflights, aes(x = dep_delay)) +
  geom_histogram(binwidth = 150)
```

# Exercise 1

1.  Look carefully at these three histograms. How do they compare? Are features revealed in one that are obscured in another?

    The three different histograms shows the different values of Bin width which can affect the appearance of the data. For instance, the second histogram shows the data falling into multiple bars while the third histogram has a large width bar and shows most of the data in one bar.

You can also filter based on multiple criteria. Suppose you are interested in flights headed to San Francisco (SFO) in February:

```{r sfo-feb-flights}
sfo_feb_flights <- nycflights %>%
  filter(dest == "SFO", month == 2)
```

# Exercise 2

2.  Create a new data frame that includes flights headed to SFO in February, and save this data frame as `sfo_feb_flights`. How many flights meet these criteria?

    The number of flights that meet this criteria is 68.

```{r}
   count(sfo_feb_flights)
```

# Exercise 3

3.  Describe the distribution of the **arrival** delays of these flights using a histogram and appropriate summary statistics. **Hint:** The summary statistics you use should depend on the shape of the distribution.

```{r}
ggplot(data = sfo_feb_flights, aes(x = arr_delay)) +
  geom_histogram(binwidth = 30, colour = "black", fill = "purple")
```

```{r summary-custom-list-origin}
sfo_feb_flights %>%
  group_by(origin) %>%
  summarise(median_dd = median(dep_delay), iqr_dd = IQR(dep_delay), n_flights = n())
```

# Exercise 4

Here, we first grouped the data by `origin` and then calculated the summary statistics.

1.  Calculate the median and interquartile range for `arr_delay`s of flights in in the `sfo_feb_flights` data frame, grouped by carrier. Which carrier has the most variable arrival delays?

    The carrier that has the most variable arrival delays is Delta (Delta) and UA(United Air Lines Inc).

Departure delays by month

Which month would you expect to have the highest average delay departing from an NYC airport?

Let's think about how you could answer this question:

-   First, calculate monthly averages for departure delays. With the new language you are learning, you could
    -   `group_by` months, then
    -   `summarise` mean departure delays.
-   Then, you could to `arrange` these average delays in `desc`ending order

```{r mean-dep-delay-months}
nycflights %>%
  group_by(month) %>%
  summarise(mean_dd = mean(dep_delay)) %>%
  arrange(desc(mean_dd))
```

# Exercise 5

5.  Suppose you really dislike departure delays and you want to schedule your travel in a month that minimizes your potential departure delay leaving NYC. One option is to choose the month with the lowest mean departure delay. Another option is to choose the month with the lowest median departure delay. What are the pros and cons of these two choices?

    The mean takes into consideration the highest and lowest numbers to come up with an average delay time while the median is the middle number and does not consider the highest and lowest values which can potentially skew the data.

# Exercise 6 

On time departure rate for NYC airports

Suppose you will be flying out of NYC and want to know which of the three major NYC airports has the best on time departure rate of departing flights. Also supposed that for you, a flight that is delayed for less than 5 minutes is basically "on time."" You consider any flight delayed for 5 minutes of more to be "delayed".

In order to determine which airport has the best on time departure rate, you can

-   first classify each flight as "on time" or "delayed",
-   then group flights by origin airport,
-   then calculate on time departure rates for each origin airport,
-   and finally arrange the airports in descending order for on time departure percentage.

Let's start with classifying each flight as "on time" or "delayed" by creating a new variable with the `mutate` function.

```{r dep-type}
nycflights <- nycflights %>%
  mutate(dep_type = ifelse(dep_delay < 5, "on time", "delayed"))
```

```{r ot-dep-rate}
nycflights %>%
  group_by(origin) %>%
  summarise(ot_dep_rate = sum(dep_type == "on time") / n()) %>%
  arrange(desc(ot_dep_rate))
```

6.  If you were selecting an airport simply based on on time departure percentage, which NYC airport would you choose to fly out of?

    La Guardia has the highest % of flights that depart on time

You can also visualize the distribution of on on time departure rate across the three airports using a segmented bar plot.

```{r viz-origin-dep-type}
ggplot(data = nycflights, aes(x = origin, fill = dep_type)) +
  geom_bar()
```

------------------------------------------------------------------------

## More Practice

7.  Mutate the data frame so that it includes a new variable that contains the average speed, `avg_speed` traveled by the plane for each flight (in mph). **Hint:** Average speed can be calculated as distance divided by number of hours of travel, and note that `air_time` is given in minutes.

```{r averagespeed}
nycflights <- nycflights %>%
  mutate(avg_speed = (distance/ (air_time/60)))
```

8.  Make a scatterplot of `avg_speed` vs. `distance`. Describe the relationship between average speed and distance. **Hint:** Use `geom_point()`.

    The average speed of flights is typically between 400 to 500 for flights going a distance of 2500 miles.

```{r viz-avg_speed}
ggplot(data = nycflights, aes(x = distance, y=avg_speed)) +
geom_point(color='blue')
```

9.  Replicate the following plot. **Hint:** The data frame plotted only contains flights from American Airlines, Delta Airlines, and United Airlines, and the points are `color`ed by `carrier`. Once you replicate the plot, determine (roughly) what the cutoff point is for departure delays where you can still expect to get to your destination on time.

    Approximately 20 minutes.

```{r plot-to-replicate, echo=FALSE, fig.show="asis", fig.width=7, fig.height=4}
dl_aa_ua <- nycflights %>%
  filter(carrier == "AA" | carrier == "DL" | carrier == "UA")
ggplot(data = dl_aa_ua, aes(x = dep_delay, y = arr_delay, color = carrier)) +
  geom_point() + xlim(-10, 50) + ylim(-10, 50)
```
