---
title: "Lab 1: Intro to R"
author: "Sanielle Worrell"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(ggplot2)
```

```{r}
data('arbuthnot', package='openintro')
```

### Exercise 1

```{r view-girls-counts}
arbuthnot$girls
```

### Exercise 2
Data shows that overall, baptism of girls were on the increase from 1629 through 1700, except for a rapid decline from 1640 through 1650. 

```{r trend-girls}
library(ggplot2)
ggplot(arbuthnot, aes(x = year, y = girls)) + geom_point(color='blue')
```


### Exercise 3

Create a new variable of boys plus girls
```{r}
arbuthnot$total <- arbuthnot$girls + arbuthnot$boys
```

Plot of boys baptized over total shows that boys accounted for over 50% of baptisms throughout the years. 
```{r plot-prop-boys-arbuthnot}
ggplot(arbuthnot, aes((x = year), y = boys / total, xlab = "Year", ylab = "Proportion of Boys Born over girls")) + geom_point(color='red') +  ylim(.5,.6)
```

### Exercise 4

bring dataset into R
```{r}
data('present', package='openintro')
```
Years included in dataset 
```{r}
factor(present$year)
```
Dimensions of the dataframe
```{r}
dim(present)
```
 Column name/variables of the dataset
```{r}
names(present)
```
### Exercise 5

Comparison of Arbuthnot data to Present shows that overall the % of boys baptized compared to boys born was low.  For example, the maximum number boys baptized is 8,426 for Arbuthnot data of total boys born 2,186,274 in the present data.

```{r}
summary(arbuthnot)
```

```{r}
summary(present)
```
### Exercise 6

Create a new variable of boys plus girls
```{r}
present$total <- present$girls + present$boys
```

Plot of the proportion of boys over the total shows that boys accounted for over 51% of births but the number of boys born compared to girls is decreasing over the years from 1980. 

```{r plot-prop-boys-present}
ggplot(present, aes((x = year), y = boys / total, xlab = "Year", ylab = "Proportion of Boys Born over girls")) + geom_point(color='red') +  ylim(.5,.6) 
```

### Exercise 7

Maximum number of births

```{r find-max-total}
max(present$total)
```
