---
title: "Assignment – Web APIs"
author: "Sanielle Worrell"
date: "2022-10-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Assignment: 

The New York Times web site provides a rich set of APIs, as described here: https://developer.nytimes.com/apis
You’ll need to start by signing up for an API key.

Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and
transform it into an R DataFrame.

```{r}
library(httr)
library(jsonlite)
library(tidyverse)
library(DT)
```


#### Get API from NY TIMES

I chosed the most popular emailed articles for the last seven (7) days
```{r}
mostemailed <- GET("https://api.nytimes.com/svc/mostpopular/v2/emailed/7.json?api-key=hG32gJN3SEk25FI2UmQkI7yBGUr6wLKH")
```

#### Convert to text file
```{r}
mostemailed <- content(mostemailed, as = "text")
```
#### Convert to dataframe
```{r}
df.mostemailed <- mostemailed %>% 
  fromJSON() %>% 
  as.data.frame()
```

#### Get column names
```{r}
colnames(df.mostemailed)
```

#### Remove certain characters in column names
Remove "." in column names

```{r}
colnames(df.mostemailed) <- gsub("[.]","_",colnames(df.mostemailed))
```

#### Select useful columns
```{r}
df_popart <- df.mostemailed %>% 
select(results_url, results_source, results_published_date, results_section, results_byline, results_title, results_abstract)
```
#### Rename column names
```{r}
df_popart <- df_popart %>%
rename("Url" = 1, "Source" = 2, "Published_date" = 3, "Section" = 4, "Author" = 5, "Title" = 6, "Summary" = 7)
datatable(df_popart, options=list(
       pageLength = 3))
```
