---
title: "Assignment - Sentimental Analysis"
author: "Sanielle Worrell"
date: "2022-11-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment:

In Text Mining with R, Chapter 2 looks at Sentiment Analysis. In this assignment, you should start by getting the primary example code from chapter 2 working in an R Markdown document. You should provide a citation to this base code. You're then asked to extend the code in two ways:

-   Work with a different corpus of your choosing, and

-   Incorporate at least one additional sentiment lexicon (possibly from another R package that you've found through research).


# Analysis

New Corpus - I chose to analyse Moby Dick, a familar tale that is known to everyone.  Code is attributed to: https://www.tidytextmining.com/sentiment.html


```{r}
# Load libraries
library(gutenbergr)
library(tidytext)
library(dplyr)
library(stringr)
library(ggplot2)
library(wordcloud)
```

```{r}
# Filter for Moby Dick
gutenberg_works() %>%
  filter(title == "Moby Dick")
```

```{r}
# Downloading title
moby_dick <- gutenberg_download(c(15),  meta_fields = "title")
```


```{r}
# Convert data using Tidy
moby_dick_chapters <- moby_dick %>%
  mutate(linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)
```

Sentimental analysis performed NRC - The NRC Emotion Lexicon is a list of English words and their associations with eight basic emotions (anger, fear, anticipation, trust, surprise, sadness, joy, and disgust) and two sentiments (negative and positive). I chosed to view all sentiments as part of this package.


```{r}
get_sentiments("nrc")
```

```{r}
nrc_sentiments <- get_sentiments("nrc")
moby_dick_sentiments <-  moby_dick_chapters %>% 
  inner_join(nrc_sentiments) %>%
  count(word, sentiment, sort = TRUE)
```


```{r}
# Plot by sentiment
ggplot(moby_dick_sentiments)+geom_bar(aes(x= sentiment, fill=sentiment))+geom_label(aes(x= sentiment, label=..count..),
            stat='count')+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust = 1))
```



```{r}
# Contribution to sentiment
moby_dick_sentiments %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)
```
```{r}
moby_dick_sentiments %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100, scale=c(1,.0005))) 
```

# Conclusion
Moby Dick has a higher count of words associated with fear next driven suprisingly by trust. The book is mostly filled with sadness. 
