---
title: "Assignment 3"
author: "Sanielle Worrell"
date: "2022-09-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

## R Markdown

#Loading data from website

```{r}
college_maj<-read.csv('https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv')
```

#1. Using the 173 majors listed in fivethirtyeight.com's College Majors dataset [<https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/>], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r}
grep(pattern = 'DATA|STATISTICS', college_maj$Major, value = TRUE, ignore.case = TRUE)
```

#2. Write code that transforms the data below:

[1] "bell pepper" "bilberry" "blackberry" "blood orange" [5] "blueberry" "cantaloupe" "chili pepper" "cloudberry"\
[9] "elderberry" "lime" "lychee" "mulberry"\
[13] "olive" "salal berry"

Into a format like this: c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

Create a vector of data

```{r}
fruits <- c("bell pepper","bilberry", "blackberry","blood orange","blueberry","cantaloupe", "chili pepper", "cloudberry",  "elderberry", "lime", "lychee" , "mulberry" , "olive", "salal berry")
fruits
```

Converting code to required format
```{r}
cat(paste0("c(",paste0(sep = '"',fruits, collapse = ', ', sep='"'),paste(")")))
```

#3 Describe, in words, what these expressions will match:

-   (.)\\1\\1
-   "(.)(.)\\\\2\\\\1"
-   (..)\\1
-   "(.).\\\\1.\\\\1"
-   "(.)(.)(.).\*\\\\3\\\\2\\\\1"

Answer:

-   (.)\\1\\1 - Any two characters that repeats in the reverse order.
-   "(.)(.)\\\\2\\\\1" - Any two characters that repeats in the same order.
-   "(.).\\1.\\1" - Any single character that repeats two more times, with each repetition after another single variable character.
-   "(.)(.)(.).\*\\3\\2\\1" - Any three characters that repeat in the reverse order after any number of variable characters.

#4 Construct regular expressions to match words that: Start and end with the same character. Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.) Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

-   Start and end with the same character.
-   Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
-   Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

Answer

```{r}
df.names <-c("anna", "church", "bob", "harry","paul", "eleven", "bubble")
regex_expr ="^(.)((.*\\1$)|\\1?$)"
str_subset(df.names,regex_expr) #keeping string matching pattern
```

```{r}
regex_expr2 = "([A-Za-z][A-Za-z]).*\\1"
str_subset(df.names,regex_expr2)
```

```{r}
regex_expr3 = "([A-Za-z]).*\\1.*\\1"
str_subset(df.names,regex_expr3 )
```
