---
title: "Project 1"
author: "Sanielle Worrell"
date: "2022-09-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyr)
library(tidyverse)

```

## R Markdown

In this project, you’re given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:

1. Player’s Name, 
2. Player’s State, 
3. Total Number of Points, 
4. Player’s Pre-Rating, 
5. and Average Pre Chess Rating of Opponents

Bringing in Chess data from Github
```{r}
Chessdata<-read.delim('https://raw.githubusercontent.com/saniewor/MSDS/main/datasets/Chessdata')
```

Removing lines with dashes by deleting every third row starting from 0
```{r}
dfChess = Chessdata %>% filter(row_number() %% 3 != 0) 
```
Renaming column
```{r}
names(dfChess)[1] <- "Chess1"
```
Selecting rows with names
```{r}
Dfnames = dfChess %>% filter(row_number() %% 2 == 1) ## Select every 2nd row starting from first row
```
Select rows with States
```{r}
Dfstates = dfChess %>% filter(row_number() %% 2 == 0) ## Select every 2nd row starting from zero
```
Split Dataframe with names (Dfnames) in column in separate columns using a delimiter
```{r}
df.Chess2 <- read.table(col.names = unlist(strsplit(Dfnames[1,], 
                                                      split = "[|]")),
                          sep = "|",
                          skip = 1, # since the headers were in row 1
                          text = unlist(Dfnames,use.names = F))
```
Split column in separate columns using a delimiter
```{r}
df.Chess3 <- read.table(col.names = unlist(strsplit(Dfstates[1,], 
                                                      split = "[|]")),
                          sep = "|",
                          skip = 1, # since the headers were in row 1
                          text = unlist(Dfstates, use.names = F))
```

Rename Columns in Chess 2
```{r}
df.Chess2 <- setNames(df.Chess2, c("Number","Playername","Totalpoints", "Rd1", "Rd2", "Rd3", "Rd4", "Rd5", "Rd6", "Rd7"))
```
Rename columns in Chess 3
```{r}
df.Chess3 <- setNames(df.Chess3, c("State","USCF ID / Rtg (Pre->Post)","No.of_games", "Rd1", "Rd2", "Rd3", "Rd4", "Rd5", "Rd6", "Rd7", "Number"))
```
Add numbers to "Number" column from blanks 
```{r}
df.Chess3$Number <- paste0(" ",1:64)
df.Chess3$Number <- as.numeric(df.Chess3$Number)
```
Separate USCF ID column into separate columns and create new dataframe with updated columns
```{r}
df.Chess4 <-separate(data =  df.Chess3, col = "USCF ID / Rtg (Pre->Post)", into = c("blank","USCF ID","R",  "PreRating", "PostRating"))
```


Merging dataframes
```{r}
df.Chessmerge<-merge(df.Chess2,df.Chess4,by.x = "Number",by.y = "Number",all=T)
```
Rearrange columns in dataframe and create cleaner dataframe
```{r}
df.Chessmerge2 = df.Chessmerge[,c(1,2,12,3,16, 4:10)] 
```
Cleaning the ratings to make them numeric
```{r}
df.Chessmerge2$PreRating <- str_extract(df.Chessmerge2$PreRating, "\\d+\\d")
```

Assigning the round columns its own variable
```{r}
rounds <- df.Chessmerge[4:10]
```
extracting the digits from the round columns, putting and putting it in a matrix
```{r}
test <- matrix(str_extract(unlist(rounds), "\\d+"), ncol = 7)
```
Creating a loop and replacing data with preratings of opponents by number
```{r}
for (row in 1:nrow(test))
{
  for (col in 1:ncol(test))
  {
    i <- test[row, col]
    
    if(!is.na(i))
    {
      test[row, col] <- df.Chessmerge2$`PreRating`[df.Chessmerge2$Number == i]
    }
  }
}
```
Changing the matrix to numeric
```{r}
class(test) <- "numeric"
```
assigning the rounds data into the dataframe
```{r}
rounds <- test
```
Calculating opponents average pre-rating, N/A values were removed
```{r}
df.Chessmerge2$Average_Opp_ratings <- round(rowMeans(rounds, na.rm = TRUE))
```
Cleaning file up to remove excess columns and generating output file
```{r}
df.Chesscleandata = df.Chessmerge2[,c(1:5, 13)] 
```
Ouput to csv
```{r}
write.csv(df.Chesscleandata, file ='Chesscleandata.csv', row.names = FALSE)
```